### 2019-12-16 doc/Manuscript.Rmd
moving notebooks to /doc folder

---
title: "Measuring the double-drift illusion with hand tracking: capturing illusion resets"
author: "Marius 't Hart"
author: "Denise Y.P. Henriques"
author: "Patrick Cavanagh"
date: "16/12/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

If a gabor pattern drifts in one direction while its internal texture drifts in the orthogonal direction, its perceived direction deviates strongly from its true direction and is instead some combination of its real external motion and its internal motion (Tse & Hsieh, 2006). The perceived location of such a double-drift stimulus periodically resets toward the veridical position. There may indeed be a simple perdiodicity underlying these resets. However, since both the veridical and perceived location are available to the brain, these resets may also occur when these two positions are too far from each other to come from the same source stimulus. Here we test which of the two explanations is more likely. In a first experiment, we had participants continuously track the position of the perceived position of the stimulus. While the orientation of the trajectories reflect stimulus properties, there were few clear resets. That’s why, in a second experiment, we had participants re-trace the perceived trajectory of a short stimulus presentation, or indicate the initial direction of the stimulus they perceived. We found resets here, and they indicate a set spatial limit to the distance between the veridical and perceived location of the double-drift stimulus.

# Introduction

Encoding the position of objects in the world is necessary for the myriad of visually guided motor tasks that we do every day. It is no surprise then that the brain has several mechanisms to gauge and update the position of objects in the world. For example, when a target is moving, its motion can influence it perceived location, making it appear to be slightly ahead of its true location (e.g. the flash lag, Nijhawan, 1994; or the flash grab effect, Cavanagh & Anstis, 2013). These cases of position extrapolation may be functional, helping the motor system overcome inevitable neural delays in targeting a moving object (Nijhawan, 1994; Duhamel et al., 1992). A stronger and very different motion-induced position shift arises when a moving gabor with internal motion orthogonal to its path is viewed in the periphery (Tse & Hsieh, 2006; Shapiro et al, 2010; Gurnsey & Baird, 2012; Kwon et al, 2015; Lisi & Cavanagh, 2015). This double-drift stimulus (Fig. 1) generates extreme misjudgments of the moving target’s location which may deviate by as much as several degrees of visual angle from its true location. Surprisingly, this extraordinary perceptual illusion does not affect eye movements to the gabor: saccades to these targets are determined by their physical not their perceived location (Lisi & Cavanagh, 2015). This result indicates that both true and illusory locations are available in the brain. Indeed, although immediate eye-movements go towards the true position, delayed eye movements and pointing go to the perceived position [Lisi & Cavanagh, 2017; Massendari et al, 2018].

The strength of the double-drift probably arises because of the very poor location information for a gabor pattern in the periphery when it has the same mean luminance as the background. With poor positional certainty, the motion information contributes to the location estimate, generating a shift in position. If the gabor itself is not moving (De Valois & De Valois, 1991), the shift saturates after about 100 ms (Chung et al., 2007) perhaps because the accumulating evidence of actual location of the stationary gabor patch is sufficient to anchor its position. However, when the gabor itself also moves as in the double-drift stimulus, there is no stable location information to anchor the position estimate and the build-up continues away from its physical location to reach deviations of several times the size of the gabor. At some point, the illusion resets, the perceived location jumps back to its physical location and then accumulation begins again (Fig. 2). This long duration before resetting is well beyond the integration period of other visual effects and into the range of short-term memory.

Here, we first examine the double-drift illusion by asking participants to track its path with a stylus on a tablet that is optically aligned with the monitor surface but with the hand not visible. Earlier results for pointing to the double-drift target (Lisi & Cavanagh, 2017) show that the illusion persists for manual pointing. However, with online tracking, the presence of the hand near the location of the perceived gabor may interact with the illusion. Specifically, the additional proprioceptive position signal may bias the perception toward its veridical location.

The ability to manually track the perceived position will also allow us to study the resets – the adjustment of perceived location when the illusion has deviated too much or too long from the physical location. It could be that the imprecise, but true, retinal position signal limits the illusion, or it could be that – like other illusions, such as bi-stable stimuli – there is a simple periodicity underlying changes in perception. Here we capture where and when observers experience resets by having them draw the perceived trajectory of the stimulus after having seen it (a retrace procedure). This will allow us to distinguish between a spatial limit to the illusion or a temporal limit.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.