### 2021-11-24 Manuscript.Rmd
revisions

---
title: "Measuring the double-drift illusion with hand tracking"
author: 
- Bernard Marius 't Hart 1,*
- Sirui Liu 2
- Peter Ulrich Tse 2
- Denise Y.P. Henriques 1
- Patrick Cavanagh 1,2,3
output:
  word_document:
    reference_docx: template.docx
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{amsmath}
bibliography: bibliography.bib
csl: journal-of-vision.csl
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# First we load all custom scripts:

# The function in this file downloads the data from the OSF project:
source('R/data.R')
# This file has some common helper functions:
source('R/common.R')
# The functions in here are used on the data of the bounded tracking experiment:
source('R/exp1_tracking.R')
# And then some code for the re-tracing experiment:
source('R/exp2_retracing.R')
# As well as models fit to that data:
source('R/exp2_models.R')

# new addition with more classic psychophysics on perceived offsets:
source('R/exp3_gaborjumps.R')

# we download the data:
OSFdata()

# we pre-process the data:

# for exp 1, this segments the tracking trajectories
# to get the middle 1 second for each full pass
segmentDirections()

# for exp 2, this extracts the reset points:
preProcessOnePass_V4()

# for exp 3, we fit psychometric functions to get PSEs:
getPSEs()

# the rest of the paper should now run without complaint
```

_1: Centre for Vision Research, York University, Toronto, ON M3J1P3, Canada_\
_2: Department of Psychological and Brain Sciences, Dartmouth College, Hanover, NJ 03755, USA_\
_3: Department of Psychology, Glendon College, Toronto, ON M4N3M6, Canada_\

*: Corresponding author: \
Marius ’t Hart\
Centre for Vision Research\
York University\
Toronto, ON M3J1P3\
Canada\
[thartbm@gmail.com](thartbm@gmail.com)\

The research was support by funding from NSERC Canada (PC and DYPH), as well as by the Department of Psychological and Brain Science, Dartmouth College (PC).

# Abstract

If a gabor pattern drifts in one direction while its internal texture drifts in the orthogonal direction, its perceived position deviates further and further away from  its true path. We first evaluated the illusion using manual tracking. Participants followed the gabor with a stylus on a drawing tablet that coincided optically with the horizontal monitor surface. Their hand and the stylus were not visible during the tracking. The magnitude of the tracking illusion corresponded closely to previous perceptual and pointing measures indicating that manual tracking is a valid measure for the illusion. This allowed us to use it in a second experiment to capture the behavior of the illusion as it eventually degrades and breaks down in single trials. Specifically, the deviation of the gabor stops accumulating at some point and either stays at a fixed offset or resets toward the veridical position. To report the perceived trajectory of the gabor, participants drew it after the gabor was removed from the monitor. Resets were detected and analyzed but their distribution neither matches a temporal nor a spatial limit on perception that could trigger resets.

# Introduction

Encoding the position of objects in the world is necessary for the myriad of visually guided motor tasks that we do every day. It is no surprise then that the brain has several mechanisms to gauge and update the position of objects in the world. For example, when a target is moving, its motion can influence its perceived location, making it appear to be slightly ahead of its true location, e.g. the flash lag [@Nijhawan1994]; or the flash grab effect [@Cavanagh2013]. These cases of position extrapolation may be functional, helping the motor system overcome inevitable neural delays in targeting a moving object [@Hogendoorn2020; @Nijhawan1994; @Duhamel1992]. A stronger and very different motion-induced position shift arises when a moving gabor has internal motion orthogonal to its path and is viewed in the periphery [@Tse2006; @Shapiro2010; @Gurnsey2012; @Kwon2015; @Lisi2015]. This double-drift stimulus (Fig. 1, left) generates extreme misjudgments of the moving target’s location which may deviate by as much as several degrees of visual angle from its true location. Surprisingly, this extraordinary perceptual illusion does not affect eye movements to the gabor: immediate saccades to these targets are determined by their physical not their perceived location [@Lisi2015]. In contrast, delayed eye movements and pointing go to the perceived position [@Lisi2017; @Massendari2018; @Ueda2018]. These findings suggest that both the true and illusory positions are available in the brain.

```{r fig1, echo=F, fig.height=3.2, fig.width=8, message=FALSE, error=FALSE, warning=FALSE, fig.cap="**Figure 1:** Stimuli and Setup **Left:** a gabor with internal motion (phase shifts) and external motion, that is viewed peripherally, appears to follow a path deviating from the true path. **Right:** participants track the perceived trajectory as the gabor moves back and forth along its path."}
library('rsvg')
library(magick)
Fig1 <- image_read_svg('doc/methods_fig_1.svg')
par(mai=c(0,0,0,0))
plot(Fig1)
```

The strength of the double-drift illusion probably arises because of the very poor location information for a gabor pattern in the periphery when it has the same mean luminance as the background [@Gurnsey2012; @Kwon2015; @Cavanagh2019]. With poor positional certainty, the motion information contributes to the location estimate, generating a shift in perceived position. If the gabor itself is not moving [@DeValois1991], the shift saturates quickly [40 _ms_: @Chung2007; 100 _ms_: @Arnold2007], perhaps because the accumulating evidence of actual location of the stationary gabor patch is sufficient to anchor its position. However, when the gabor itself also moves, as in the double-drift stimulus, there is no stable location information to anchor the position estimate and the build-up continues well beyond 100 ms. The perceived location continues to move away from its physical position by up to several times the size of the gabor.

The strength of the illusion; the deviation of perceived path of the gabor from its true path, depends on the speed of both the internal and external motion [@Tse2006; @Shapiro2010; @Kwon2015; @Cavanagh2019; @Heller2021]. Rather than use point estimates of the perceived location of the stimulus, we want to test how well a continuous trajectory can measure the strength of the illusion. Eye and hand movement are differentially sensitive to many illusions, such as the Ponzo illusion [@Gamble2017], Ebbinghaus illusion [@Knol2017], flash-lag effect [@Blohm2003], the Duncker illusion [@Soechting2001] and many others. Since it has been shown that the double-drift illusion affects pointing movements, but not saccades [@Lisi2017] we expect that hand trajectories will also be sensitive to the illusion.

We assess this in our first experiment, using hand movement trajectores recorded during stimulus presentation (Fig. 1). We compare trajectories across 5 different internal motions of the gabor and compare the illusion strength with previous work based on a straightforward vector combination model [@Heller2021]. As we expected, online tracking provides a similar measure of illusion strength as pointing.

In our second experiment, we look at the spatial and temporal limits of the illusion. At some point, the illusion stops accumulating. The accumulation can be reset by introducing a temporal break [@Lisi2015] or by distracting attention [@Nakayama2020], but it may also reset spontaneously once the accumulation has gone on too long or too far. Informally, some observers have reported that the gabor’s path may saturate so that it remains parallel to, but offset from the true path. Others report that the position moves back toward its physical location, either slowly or suddenly, whereupon accumulation begins again. The ability to record trajectories of the perceived position will then allow us to study spontaneous reset outcomes.

# Experiment 1: Online tracking

In this experiment, participants were asked to use a stylus on a drawing tablet to track a moving gabor with some amount of internal motion that elicited the double-drift illusion. A variant on online tracking has been used before [@Cormack2019] with a continuous task where the stimulus depended on the response. We test if tracking the illusion continuously is a good measure of illusion strength, or if it prevents or decreases the illusion.

Potential extra citation: [@PatricioDecima2022].

## Methods

### Participants

For this experiment, 4 participants were recruited from the lab (3 female; ages 21 – 31, mean: 25.75). All participants reported being right handed and had either normal or corrected to normal vision. Procedures were in accordance with the Declaration of Helsinki and were approved by York’s Human Participants Review Committee. All participants provided prior, written, informed consent.

### Setup

Participants used a stylus on a horizontal drawing tablet (Wacom Intuos Pro) to indicate where they perceived the location of a gabor (see Fig. 1, right), as well as a keypad for additional responses. Stimuli were displayed on a downward facing LCD monitor (30 Hz, 20”, 1680x1050), parallel with the drawing tablet. The stimuli were observed via an upward facing mirror placed exactly between the tablet and monitor, so that the stimuli appear to be in the same horizontal plane as the hand, and the tip of the stylus. Experiments were run in Python 2.7 with PsychoPy [@Peirce2019].

For reporting degrees of visual angle that stimuli spanned ( _dva_ ), we assumed the participants to be close to the setup, with their eyes at the height of the monitor. While there was no requirement to use the forehead rest, this is a reasonable assumption, since stimuli are only visible when relatively close to the setup. Nevertheless, reported _dva's_ are estimates of the size of stimuli.

### Stimuli

The stimuli were phase-changing gabor patches with a sigma (sd) of 0.43 _cm_ or ~0.59 _dva_, spatial frequency of 0.58 _c/cm_ or ~0.42 _c/dva_, vertical orientation, and Michelson contrast of 98.7%. The background luminance was 18.46 $cd/m^2$ and the mean luminance of the gabor deviated from the background by less than 1%. By simultaneously moving the gabor envelope and drifting its internal sine wave, it becomes a ‘double-drift stimulus’ where the perceived path deviates from the real path. The gabor changed phase at different rates, called 'internal motion' here, specified in cycles per second ( _cps_ ). The real displacements of the stimulus are along the Y axis of the monitor, keeping the X coordinate constant at the middle of the screen, so that a stimulus without internal motion would appear to move toward and away from the participant on the horizontal display.

The motion of the envelope was restricted to the central 13.5 _cm_ ( ~14.0 _dva_ ) and was set so that the stimulus would move from one end to the other of its path ( a motion we call a 'pass' ) in 2 _s_, corresponding to a speed of ~6.75 _cm/s_ ( or roughly 7.0 _dva/s_ ).

### Procedure

In this experiment, participants used their unseen hand and stylus to do continuous online tracking of the double-drift stimulus while that moved back and forth along the Y-axis of the horizontal screen for 12 _s_. Before trial onset, the participant moved the stylus to the middle of the screen, and were given 1.5 _s_ to fixate a point on the left or right of the screen. The gabor would then appear at the middle of the screen and start moving. Internal motion could be 3, 1, 0, -1 or -3 _cps_. Since each moving gabor was shown for 12 _s_, and it moved from one end to another in _2_ s, it completed 5 full and 2 half passes of the path. The internal motion of the gabor was inverted at the far and close ends of the physical path, where external motion was also inverted.

### Analyses

Because there is no set reference point to gauge movement direction against after each direction change, we use instantaneous heading along the trajectory instead. We first segmented the tracking trajectories according to full stimulus passes (in between direction changes of the gabor), and removed the noisy tracking during the first and last half second, leaving 1 _s_ trajectory segments. We then calculated the instantaneous heading (direction, disregarding velocity or distance) between all 32 raw samples for 31 heading samples per segment. This heading measure should depend on the illusion strength.

We compare the illusory direction of the heading against the predictions of a simple vector combination model of perceived direction from Tse and Hsieh [-@Tse2006], Cavanagh and Tse [-@Cavanagh2019] and Heller et al. [-@Heller2021]. In this model, the perceived direction is a simple vector combination of the external ($V_e$) and internal ($V_i$) motions and a constant $K$. That is, the deviation from the physical direction is given by:

[1]  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  $\theta=tan^{-1}K\frac{V_i}{V_e}$

Here we calculate $V_i$ and $V_e$ in _cm/s_; _cps_ * 0.58 _cm_ for $V_i$ and 13.5/2 _cm/s_ for $V_e$. If $K$ is 0, then there is no illusion and the perceived direction matches the external motion direction. However, if $K$ is 1, then the internal motion has equal contribution with the external motion in determining the perceived direction. For comparison with other work, we will find the value of $K$ that best describes the strength of the illusion in each of our two experiments.

In particular, we compare the illusion strengths we find with those reported by Heller et al. [-@Heller2021]. They show a double-drift stimulus with 36 combinations of internal and external motion for 500 ms and then have participants indicate on a ring where they perceived the path of the stimulus to intersect with this ring. While they hypothesize that illusion strength decreases with faster external motion (perhaps because of resets or limits on perception), here we will mostly use their overall estimate of illusion strength, given by $K=0.74$.

All analyses were done in R 3.6.1 [@Rproject]. All data, as well as scripts for the experiments and analyses are available on OSF: [https://osf.io/72ndu/](https://osf.io/72ndu/).

## Results

All raw trajectories are show in Figure 2A. In Figure 2C, we plot the average instantaneous heading during the middle second of each 2 _s_ pass for each of the 5 internal motion speeds. The average angle of the tracked path is given in degrees deviation from straight forward. It seems that continuous tracking is sensitive to the strength of the illusion, and this is confirmed by a repeated-measures ANOVA on the average angles, using internal motion as a within-subjects factor ( _F_(4,12)=102.6, _p_<.001, $\eta^2$=0.97 ). With 4 FDR-corrected [@Benjamini1995], paired t-tests, we find a difference in heading between internal motion speeds of 0 _cps_ and both 1 _cps_ ( _p_=.022) and -1 _cps_ ( _p_=.025) as well as between -3 and -1 _cps_ and 3 and 1 _cps_ (both _p_ <.001). This makes it clear that continuous tracking does not prevent the illusion from occurring. We will use the heading angle as a measure of the strength of the illusion.

```{r fig2, echo=FALSE, message=FALSE, fig.width=6, fig.height=5, fig.cap="\\label{fig:fig2}**Figure 2:** Online Tracking Data. **A:** Raw individual trajectories. In general, the traced paths approximate linearity, with the direction depending on the internal motion. **B:** Four trials with a possible reset. **C:** Average heading of the middle 1 _s_ of each pass. Distance from origin is time, angle is the heading angle. Dark blue lines show the average heading over time. The light blue depicts the average area (across participants) containing 90% of the heading data. The darker blue areas depict the 95% confidence interval of the mean heading. **D:** The strength of the illusion in degrees angle deviation from the physical path. In red are the average illusion strengths for 0 _cps_, ±1 _cps_ and ±3 _cps_, and in light blue the data for the 4 individual participants. Black line: prediction of the vector combination model of Eq. 1 with $K=0.74$ as in Heller et al. (2021), which is indistinguishable from the best fit to our data: $K=0.738$."}
plotBoundedTracking()
```

## Illusion strength

The data clearly show the illusion, but does tracking decrease the strength of the illusion compared to simply observing it? In Figure 2D, we compare mean illusion strength to that reported previously [@Heller2021]. As can be seen the averages from our data here (red circles) coincide fairly well with the illusion strength predicted by the model fitted to all their data (Eq. 1, black line). If we fit the vector combination model, we get $K=0.738$ which is very close to $K=0.74$ reported by Heller et al. [-@Heller2021]. That is; tracking the illusion while observing it does not change the strength of the illusion.

# Experiment 2: Delayed retracing

The illusion cannot drift continuously away from the physical path forever. It can be disrupted by a temporal break [@Lisi2015] or by distracting attention [@Nakayama2020]. These disruptions may cause the perceived path to stay at a fixed offset, travelling parallel to the physical path, or it may return toward the veridical position before resuming the illusory direction (Fig. 3). Other percepts of resets have been reported as well, but as far as we know these fall in between these two extremes. Resets may also occur spontaneously once the accumulation has gone on too long or too far, depending on what mechanism causes resets. Where resets occur would then both depend on the mechanism that causes resets, as well as on the strength of the illusion (purple arrows in Fig. 3) which in turn depends on the stimulus properties as well as variation in perception within and between participants. The purpose of this experiment is to capture this spontaneous resetting and determine its source.

In the previous experiment with online tracking, the duration of each pass in the online tracking experiment was 2 seconds and the manual traces (Fig. 2A,B) mostly showed a linear trajectory without saturation or reset of the types shown in Figure 3 (left). The continuation of the accumulation may be a result of the active tracking – either the extra attention required to track increased the extent of the accumulation or possibly the temporal demands of the tracking caused the participants to average their traces over any breaks. Alternatively, post-dictive inference may prevent resets from being perceived. To address this, we switched to a delayed, offline recording of the perceived path.

```{r fig3, echo=F, fig.height=3.4125, fig.width=6, message=FALSE, error=FALSE, warning=FALSE, fig.cap="\\label{fig:fig3}**Figure 3:** Double-Drift Resets. **Left:** At some point, the perceived position may stop moving further away from its actual path. For some people this takes the form of a ‘return’ reset, back toward the true position either suddenly or slowly, for others it means the illusory position remains at a fixed offset moving parallel to the real path. **Right:** Given varying strength of the illusion (denoted by the angle of the purple arrows), e.g. due to stimulus properties or within or between subject differences, resets may occur at different points in space. **Top right:** if these resets can be explained by a spatial limit on the size of the illusory position shift away from the true vertical path, the locations of the resets should have a constant X coordinate; the spatial offset from the vertical path. **Bottom right:** if they occur instead after some period of time, resets should have a constant radius (R); the time since the gabor’s motion began."}
library(magick)
Fig3 <- image_read_svg('doc/methods_fig_2.svg')
par(mai=c(0,0,0,0))
plot(Fig3)
```

We were interested, in particular, in whether the limitation of the accumulation would be set by space or time. Since both the true and the illusory position appear to be available in the visual system (the true position drives saccades whereas the illusory position drives perception, @Lisi2015), it is possible that resets occur once the distance between the real and perceived positions exceeds some limit. We can speculate that this spatial limit would be related to the positional uncertainty of the gabor. When the position information is reliable, the resets would occur with very little illusory deviation, keeping the perceived path close to the true path. When the positional uncertainty is high, in the periphery, with the gabor’s mean luminance matched to the background [@Gurnsey2012; @Cavanagh2019], the spatial offset could be quite large before exceeding the range of positional uncertainty around the gabor’s true position. Under this hypothesis, the occurrence of spontaneous resets will depend on the spatial offset from the true path which will be directly proportional to the internal speed (for a constant external motion). On the other hand, if spontaneous resets are the result of the temporal limitations of the integration process, resets would occur after a certain amount of time, independently of the internal speed and the spatial offset it creates. Investigating the time and location of resets in the recorded trajectories, allows us to distinguish between a spatial limit to the illusion and a temporal limit. This in turn will inform us of the processes underlying the perception of the position of moving objects.

To do so, we ask participants to re-trace the perceived path of a double-drift stimulus, immediately after viewing the stimulus, a method used by Nakayama & Holcombe [-@Nakayama2020]. This allows participants more time to carefully reproduce the path. We use the recorded trajectories to assess points where resets occurred, whether they were a discontinuity in orientation reflecting a saturation (hit the wall reset, Fig. 3) or a discontinuity in position reflecting a jump back to the real location. These reset points can then be used to determine if resets are time-limited or space-limited.

## Methods

### Participants

For this experiment, 9 participants were recruited from the lab (6 female; ages 19 - 27, mean: 22.8). All participants reported to be right handed and had either normal or corrected to normal vision. Procedures were in accordance with the Declaration of Helsinki (2003) and were approved by York’s Human Participants Review Committee. All participants provided prior, written, informed consent.

### Setup and Stimuli

The setup was the same as used in the online tracking experiment, except that we now also used a small keypad for responses (Fig. 1, right). Stimuli were identical to those used in the online tracking experiment, but now always started at the near end of the workspace and moved away from the participants. The double drift stimuli would use either 3 or 4 seconds to move 13.5 _cm_, corresponding to speeds of ~4.5 and ~3.75 _cm/s_ or ~4.7 and ~3.5 _dva/s_, and had an internal drift of 2, 3 or 4 _cps_ .

## Procedure

In experiment 2, participants first watched a double-drift stimulus and then either reproduced the perceived path of the stimulus by retracing it on the tablet, or indicated the initial movement direction by changing the orientation of a line originating in the same position as the doube-drift stimulus (data not used, but corresponds to illusion strength determined from trajectories). More details are given below. This provided more time for the participants to reproduce any resets, i.e. without the need for real-time tracking. The re-tracing task was done in half of 8 blocks (the other half of the blocks are not used). Block types were alternated, and the order was counterbalanced across participants. Each block used the 6 combinations of two external speeds ( corresponding to 3 or 4 _s_ presentation time ), and three internal speeds ( 2, 3 or 4 _cps_ ), 6 times, for a total of 36 trials per block. Each of the combinations of internal and external speed was presented 24 times in total, so that we have 144 trials for each participant, and 1296 trials in total.

In both kinds of trials, participants first had to move the stylus to the start position of the gabor, and then fixate a point to the left or right of the centre of the workspace. The double-drift gabor started at the near end of the display, and moved along the horizontal display away from the participant, always making a single, forward pass. After the gabor had disappeared, participants could respond in one of two ways. When re-tracing the perceived path with the stylus on the tablet, the drawn path would show up as a red line. Participants could "key in" their response by pressing ENTER, or start over by pressing ESCAPE. There was no time-limit for the response. Re-tracing the perceived paths should allow capturing spontaneous resets of the illusion, if there are any.

### Analyses

The individual trajectories reveal some points that are potential spontaneous resets that we try to extract. Using a smooth spline interpolated version of the trajectory we detect right-left direction changes, based on sign changes. First, we remove the first and last 4 mm of the trajectory, as well as any short segments on the trajectory at the start and end that go toward the participant, to remove any unintended jitters caused by starting and ending the movement. Then, we apply a bi-directional low-pass Butterworth filter with a cut-off frequency of 1.5 Hz. However, since participants draw at very different speeds, we first linearly interpolate the trajectories such that the sampling rate is normalized to 30 Hz for all trajectories, based on the stimulus duration. In these summary trajectories we detect the first inflection point using the X coordinates. This could put the reset point too late for hit-the-wall resets, so that we move back to get the previous sample at 95% of the X coordinate, and finally locate the sample in the raw trajectory that is closest to this point. Finally, we remove a few reset points that are detected very close to the veridical path without any build-up of the illusion by only keeping reset points that fell 5 mm within the quadrant where we expect them. This heuristic will result in some misses and some false alarms, and while changing the algorithm may improve performance, there is no ground truth to evaluate this. All raw trajectories and the detected reset points can be found on the OSF repository (XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX add link?).

To facilitate modeling, the horizontal and vertical (straight ahead) coordinates of the reset points are kept isometric by giving them in centimeters. The distance from origin can then be converted to time as 1 _cm_ = 4/13.5 _s_ in the 4 _s_ condition and 1 _cm_ = 3/13.5 _s_ in the 3 _s_ condition.

Apart from reset points, we need a measure of illusion strength. We take a trajectory sample halfway between the starting point and the detected reset point. We then use the angular difference between the gabor's real trajectory and a straight line drawn through that point and the start of the trajectory as a measure of illusion strength (see Fig. 4).

## Results

In Figure 4 (middle), we show the trajectories for participant 4 in the condition with 3 _cps_ internal motion and 4 _s_ passes. This particular participant's re-traced paths show clear 'return' resets, all of which are detected by our simple algorithm. Other participants' data are less clear.

```{r fig4, echo=FALSE, message=FALSE, fig.width=6, fig.height=6, fig.cap="\\label{fig:fig4}**Figure 4:** Re-tracing Data. **A-D:** Example trajectories and reset points. All trajectory samples in transparent light blue (denser samples result in darker color). Dark red lines are, pruned, interpolated and filtered trajectories (see Methods). Purple circles indicate reset points. **A:** Trajectory with several sudden resets. **B:** 'Hit-the-wall' reset. **C:** Slow return reset. **D:** Continued accumulation of the illusion. **E:** Distribution of all 827 detected reset points. **F:** Using the angle to a point along the trajectory at half the distance to the reset point as a measure of illusion strength, we can compare this data to other data. The illusion is weaker ($K=0.60$, blue line) and its strength varies between participants, but it is nevertheless clearly there."}
#plotExampleData()
plotResetPoints()
```

On average, this method detects resets in ~68% of trials. For two participants, less than half the trials showed resets (9% and 22%). For the remaining seven participants the algorithm detected resets in 65%-99% of trials, with an average of 82%. Since participants experiencing a "hit-the-wall reset" (Fig. 3) as opposed to a "jump" reset, might not always display a detectable direction change in their paths, most of these percentages of trials with resets seems reasonable. However, the participant with 9% resets will not be used for figures and statistics, but we have verified that excluding their data does not change the results. This means we analyse a total of 858 reset points (Fig. 4, right).

As before, we compare the strength of the illusion (see Fig 5, left), with an earlier study [@Heller2021]. A line through the origin with $K=0.54$ predicts illusion strength fairly well in these data, which is weaker than the $K=0.74$ found by Heller et al. [-@Heller2021]. Illusion strength is also lower than in Experiment 1 with online tracking and it has a fair amount of between-subject variation in this task (blue shaded area is 95% confidence interval across participants). While the illusion is less strong, it is nevertheless there and depends on the relationship between internal and external speed as in previous studies.

## Limits on the Double-Drift Illusion

Now we look at the distribution of reset points, and whether they can be explained by a temporal or spatial limit. In Figure 4E, we can see that the reset points vary more along the y-axis, than along the x-axis. This would be consistent with a spatial limit to the illusion resulting in resets. However, there is also considerable variance, so that fitting any line through the data, will necessarily leave errors.

```{r fig5, echo=FALSE, message=FALSE, fig.width=6, fig.height=5.333333, fig.cap="\\label{fig:fig5}**Figure 5:** Reset models. **A:** Model limits relative to reset point locations. Reset points neither fall along a spatial limit (blue line) or a temporal limit (red lines). Marginal distributions show that reset points' X coordinates and time since stimulus approximately fall in a gamma distribution. **B:** Gamma distribution on time (log likelihood: -766.6). **C:** Gamma distribution on spatial offset (log likelihood: -1319.3)."}
#plotResetPoints()
invisible(capture.output(plotModels()))
#plotBootstrappedFits()
#plotIllusionStrength()
```

We test this with a series of simple models. First, for every reset point (with coordinates $X$ and $Y$) we use the illusion strength based on a point earlier in the trajectory, expressed as either an angle (Fig. 4, left: $a = \arctan{(X/Y)}$) or a slope ($s = \sin{(a)}/\cos{(a)}$) to predict reset points.

If the perceived path of the gabor is approximated by a straight line (up to a reset), and reset points occur at a set limit, then a spatial limit ($L_x$) would equal the x-coordinate and predict the y-coordinate, $y$, of the first reset:

[2]  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  $y = L_x \cdot s$

And a temporal limit ($L_t$) would predict both coordinates of the reset, $x$ and $y$:

[3a]  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  $x = L_t \cdot v \cdot \cos{(a)}$

[3b]  &nbsp;&nbsp;&nbsp;&nbsp;&nbsp;  $y = L_t \cdot v \cdot \sin{(a)}$

Where $L_t$ and $L_x$ are a temporal and spatial limit (in seconds and cm respectively) that are free parameters, $a$ and $s$ correspond to the calculated angle and slope expressing the strength of the illusion, and speed ($v$ in _cm/s_) is given by the stimulus. If we fit the two simple models for each participant separately, by minimizing the mean square error (in isometric space, with absolute distance in centimeters between predicted and actual reset points), most participants are best fit with a spatial limit (MSE: 6/9, AIC: 7/9). This indicates that in this data, a spatial limit explains the reset points best, but neither limit explains the data very well.

To obtain population level values for the spatial and temporal limits, we resample reset points from each participant, using the smallest number of detected reset points that fall in the first quadrant, in the group (44), to ensure that each participant contributes the same number of reset points. Both models based on a spatial limit (eq. 2) and on a temporal limit (eq. 3)  were fit 1000 times (Fig. 5, right). The median fits use a spatial limit of 3.5 _cm_ or ~3.6 _dva_ and a temporal limit of 1.1 _s_. We explored other modeling approaches (e.g. based on various distributions of likelihoods or combining limits) but these show the same pattern of results: spatial limits are somewhat favoured, but temporal limits can not be excluded (see OSF repository: [https://osf.io/72ndu/](https://osf.io/72ndu/)).

# Discussion

We first tested whether manual tracking of a double-drift stimulus is susceptible to the illusion as reported by perceptual judgments, pointing [@Lisi2017], and memory saccades [@Massendari2018; @Ueda2018]. The results of this online tracking experiment show a close agreement between the angle of the manual tracking and the illusion strength expected based on the model from earlier perceptual measures [@Cavanagh2019]. However, since no unambiguous resets were captured in the traces of this experiment, we followed up with a second experiment that did record evidence of resets.

To measure resets, we asked participants to draw the perceived path of the stimulus after a single pass. The orientation of the drawn traces reflect the strength of the illusion, and in most trials there was a clear reset point, for either a "return" reset or a "hit-the-wall" reset. We then set out to test if these spontaneous resets are triggered by a limiting distance of drift away from the physical location, or by a specific limit of time. The results showed that the resets are slightly better explained by the distance from the true position, but a temporal limit is not ruled out.

Recently, Kwon, Tadin, and Knill [-@Kwon2015] proposed a Bayesian object tracking model to explain the double-drift illusion. In their model, when the position signals are weak, the motion signals bias perceived position away from the true location. In their case, this deviation saturated after about 200 _ms_ and the perceived path then continued parallel to the physical path, as in the “hit the wall” resets we described above [@Kwon2015, Fig. 3]. In contrast to their model, our participants in Experiment 2 showed quite variable timing for the reset point where the accumulation of offset ended. The resets were more stable in terms of spatial offset (Fig. 4, middle) but much more variable in terms of time from the start of the gabor's motion. In any case, the temporal limit of these reset points was not 200 _ms_ but 1.1 _s_ and fell into a wide range from 0.5 _s_ to almost 4 _s_ across trials (see Fig. 4). Moreover, 4/9 of our participants regularly reported “return” resets that the Kwon et al. [-@Kwon2015] model does not predict. The “return” resets suggest that both the physical and perceived locations are available to the visual system and that when the deviation exceeds some limit, the perceived location returns toward the physical location. At that point, the build-up of the illusion may start again.

Recently, it has been suggested that resets can be triggered by a distraction of attention [@Nakayama2020]. The participants in this experiment also retraced the perceived path and reported “return” resets. Nakayama and Holcombe also suggest that planned eye movements can trigger a “return” reset back to the physical path. In our experiment, we did not control for eye-movements, so we can't directly investigate whether eye movements accompanied the “return” resets that participants reported. However, the modestly better performance of the spatial limit model here suggests that spontaneous resets are not randomly distributed - as they might be if eye-movements or shifts of attention are the only way to elicit resets.

We find that manual tracking of the double-drift stimulus during its motion did show the expected illusion. Few if any resets were detected during the online tracking, either because of the additional attentional requirements or because of the short duration (2 s). Participants did report spontaneous resets in delayed, single-pass re-tracing of Experiment 2. The resets were best explained by a limiting spatial offset from the true stimulus location, but do not exclude a temporal limit. These findings may either indicate there are multiple causes for spontaneous resets or that a re-tracing method is not suitable to resolve what causes this phenomenon. Either way, both some amount of time from stimulus onset until reset and too large a conflict between retinal and perceived position leading to resets remain possible.

# Acknowledgements

The research was support by funding from NSERC Canada (PC and DYPH), as well as by the Department of Psychological and Brain Science, Dartmouth College (PC).

# References {#references .unnumbered}