### 2021-12-16 DoubleDriftResetLimits.Rmd
deleted everything from the notebook that is no longer relevant (most of it) will have re-write this document later

---
title: "Double-Drift Illusion Boundaries"
output:
  html_notebook: default
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

# Overview

We want to see if we can find reset points in tajectories people draw while or after seeing a double-drift stimulus, with the goal to see if we can figure out if those resets occur after a certain time interval (like perception in other illusions) or if it occurs when the percept reaches certain spatial difference with the true position.

## Source files

There are several files with functions that will do the work for us in pre-processing, plotting and doing stats.

```{r}
# The functions here allow downloading the collected data from the OSF project:
source('R/data.R')
# This file has some common helper functions:
source('R/common.R')
# The functions in here are used on the data of the bounded tracking experiments:
source('R/exp1_tracking.R')
# And finally some code for the re-tracing experiment...
source('R/exp2_retracing.R')
# As well as models fit to that data:
source('R/exp2_models.R')
# And to make comparisons with the Liu et al. (preprint) data:
source('R/exp3_gaborjumps.R')
```

## Download data

A last step before starting actual analyses is to download the data from the Open Science Framework:

```{r}
# from R/data.R
OSFdata()
```

If there were no errors, we're now ready to start.

# Experiment 1a: continuous tracking

In this experiment, 4 participants were instructed to look at a fixation point at the left or right side of the screen, while moving a tablet controlled cursor to the center of the screen. Subsequently, a double-drift stimulus would appear and then move on a straight, 13.5 cm path along the Y axis of the screen (not sideways, but forward/backward, or upward/downward if you like), starting in one of two directions at random. When the double-drift stimulus would reach either end of the path, both it's external and internal motion would be inverted. This way it would appear to move on a diagonal, with the angle of the diagonal depending on the direction and velocity of the internal motion. Participants were to use the stylus on the tablet to track the perceived position of the stimulus while it was moving. The stimulus would complete 3 full cycles (if one path is the 13.5 cm: a half path, five full paths, and another half path) in 12 seconds. That is: a single path takes 2 seconds. The internal motion could have one of 5 values -3, -1, 0, 1 or 3 cycles per second (the number of complete phases the gabor would cycle through in 1 second). Each of these 5 internal motion values was repeated 8 times (with random initial direction and fixation side) for a total of 40 trials per participant.

First, we process the raw data into a more useful format for our analyses:

```{r}
# is there a function for this?
```

This will (re?) create pre-processed data files in the directory: `data/bounded_tracking`.

Then we plot the trajectories recorded during this task (each subplot essentially shows 8 trials overlayed), and in the top row we show the average heading as a thicker line, as well as a 2D polar histogram showing the distribution of headings.

```{r fig.width=6, fig.height=6}
# from R/tracking_bounded.R:
plotBoundedTracking()
```

```{r}
analyzeBoundedTracking()
```

A lot of stuff is missing right now. Will fill in later.

## Expected resets in experiment 1, based on numbers in experiment 2

If we expect resets at the same rate as they occur in experiment 2, and each pass in experiment 1 is independent. This becomes a fairly simply calculation of chance.

```{r}
# data:
df <- getDataTrials()

# percentage of trials with resets before 1 second (0.2229938)
sp <- sum(df$RT < 1) / 1296
# percentage of trials with resets before 2 seconds (0.5223765)
lp <- sum(df$RT < 2) / 1296

exp1trial <- 1 - ( (1-sp)^2 * (1-lp)^5 )
exp1trial

```

This would mean that 98.5% of trials in exp 1 should have shown at least 1 reset. Let's see if this is statistically different from the 6/160 (~4%) that we observed:

```{r}
binom.test(6, 160, p=exp1trial)
```

Of course it is.

## Figures for Journal of Vision

From the website it is not directly clear which formats are and are not accepted, but it seems they want vector-based formats like pdf or eps. Let's go with pdf for now.

```{r}
replotMethodsFigures(target='pdf', figno=1)
plotBoundedTracking(target='pdf')
replotMethodsFigures(target='pdf', figno=3)
plotResetPoints(target='pdf')
plotModels(target='pdf')
plotJumpPSEs(target='pdf', individualDots=F, modelDots=F, gammaDots=F, modelLine=T, gammaLine=T, propExpect=T)
```